---
title: "Project_241_DK"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data Analysis

In the code chunk below we combine our datasets i.e all the emails that we have sent in 5 pushes in a single data.

```{r Combining the data}
#reading the file group1


our_emails = c("mirza2020@berkeley.edu", "deveshkhandelwal@berkeley.edu",
               "rahosbach@berkeley.edu", "tgao2020@berkeley.edu")

#reading in the data for 02nd July

dg1 <- fread('data/round1_0702_group1_0702_0941.csv')
head(dg1)
#retaining the relevant columns
dg1 <- dg1[,c('Email','Merge status')]
#verifying
head(dg1)

#reading the file group2 for 2nd July
dg2 <- fread('data/round1_0702_group2_0702_1028.csv')
head(dg2)
dg2 <- dg2[,c('Email','Merge status')]
head(dg2)

#reading the file group3  for 2nd July
dg3 <- fread('data/round1_0702_group3_0702_1055.csv')
head(dg3)
dg3 <- dg3[,c('Email','Merge status')]
head(dg3)

#combined data for round1
round1 <- rbind(dg1,dg2,dg3)

print(round1[,.N])

#reading the data for round 2 - Round 2 onwards we used a single spreadsheet for the ease of tracking.
round2 <- fread('data/round2_0709_all_0716_1117.csv')
head(round2)
#retaining the relevant columns
round2 <- round2[,c('Email','Merge status')]
#verifying
head(round2)


print(round2[,.N])

#reading the data for round 3 - This is combined
round3 <- fread('data/round3_0716_all_0722_1046.csv')
head(round3)
#retaining the relevant columns
round3 <- round3[,c('Email','Merge status')]
#verifying
head(round3)


print(round3[,.N])

#reading the data for round 4 - This is combined
round4 <- fread('data/round4_0722_all_0723_1131.csv')
head(round4)
#retaining the relevant columns
round4 <- round4[,c('Email','Merge status')]
#verifying
head(round4)


print(round4[,.N])

#reading the data for round 4 - This is combined
round5 <- fread('data/round5_0723_all_0724_1219.csv')
head(round5)
#retaining the relevant columns
round5 <- round5[,c('Email','Merge status')]
#verifying
head(round5)


print(round5[,.N])


```
We have created multiple datatables, the cde chunk below will concatenate the data horizontally.
We also remove our group's email , which was added to ensure delivery with a mail merge techbology.
```{r combining all the files}

d_combined <- rbind(round1,round2,round3,round4,round5) %>% filter(!Email %in% our_emails)

head(d_combined     )


#sum(d_combined[, Email == "deveshkhandelwal@berkeley.edu"])
```
We now analyze the aggregated data.
The idea is to check the status of the emails by looking at the column `Merge Status`

```{r checking the status by email }
library(dplyr)
#https://stackoverflow.com/questions/25571547/select-unique-values-with-select-function-in-dplyr-library
uniq_status <- d_combined %>% distinct(`Merge status`)

uniq_status

# not sure what the 0 status is
d_combined[ ,.N]

d_combined[`Merge status`== 0 , Email]

d_zero_status <- d_combined[`Merge status`== 0 ,]


```
We notice that there are certain emails with status as `0`. some of these emails overlap with the bounced emails but nonoe of the people with the emails in this category responded to our survey. We therefore remove these `r  d_zero_status[,.N] `emails from further analysis.
```{r removing merge status zero}
#storing the data 
write.csv(d_zero_status,"merge_status_0.csv",row.names = TRUE)
#getting rid of the 105 rows with zero merge status
d_combined_non_zero <- d_combined[`Merge status`!= 0 ,]

d_zero_status[,.N]
```

## Finding Compliers
Compliers are the people in our dataset who have at-least opened an email once.We have sent the emails out in 5 pushes and a person with a valid email id would have received more than a single email.
In the code chunk that follows we calculate the number of compliers.
Our strategy is to group the data by emails and summarize the groups by emails when there is at least one read email.


The function below combines the emails and adds a new column in the dataframe with status TRUE if the person has read the email at least once.
```{r function definition }

final_mail_status <- function(d , status)
  # This function takes the combined data frame and conslosidates the dataframe rows on email status
{
  
  #first we'll decode the status we want to convert as TRUE
  if ("Merge status" %in%  colnames(d))
    {
    #delete column
    d <- subset(d , select = c("Email","Merge status"))
    }
  
  
  
  d_merged <- setDT(d)[`Merge status`== status ,Email_status:= (as.logical(1))]
  #everything else to 0
  #d_merged <- setDT(d)[is.na(`Email_status`) ,`Email_status`:= (as.logical(0))]
  library(dplyr)
  
  #d3 <- d2 %>%  group_by(Email) %>% summarise(status = any(as.logical(`Merge Status`)))
  d_summarized<- d_merged %>% group_by(Email) %>% summarise(status = any(as.logical(`Email_status`)))
  
  return (d_summarized)

}

```


## Data Consolidation

We now focus our attention to the combined dataset, the idea is to assign a status to each of the unique email id in our alumni corpus and determine whether that person is a complier , attritor or Never-taker.

```{r finding open emails}

#calling function to get the list of those who have opened email at least once
d_at_least_one_read <- final_mail_status(d_combined_non_zero, "EMAIL_OPENED")



#This has NAs too that could correspond to other status, Lets remove the NAs.
d_at_least_once_open <- d_at_least_one_read [!is.na(d_at_least_one_read$status),]

#saving the data to a file
write.csv(d_at_least_one_read,"email_opened.csv",row.names = TRUE)

#getting the details for those who have clicked the emails.

d_at_least_one_clicked <- final_mail_status(d_combined_non_zero, "EMAIL_CLICKED")
d_clicked <- d_at_least_one_read [!is.na(d_at_least_one_clicked$status),]

write.csv(d_at_least_one_clicked,"email_clicked.csv",row.names = TRUE)


d_at_least_one_bounce <- final_mail_status(d_combined_non_zero, "EMAIL_BOUNCED")
sum(d_at_least_one_bounce$status == TRUE)



```

# Data Aggregation - Joins
We now combine our data from our alumni scrape and assign a `final status` to each email.



```{r JOINS}
#reading the alumni file
alum_data <- fread('data/alumni_scrape.csv')

#reading in the open emails
open_data <- fread('data/email_opened.csv')
#chaaging the names for join to work
setnames(alum_data, "email", "Email")

print(colnames(alum_data))
print(colnames(open_data))

#performing left join - We want all the cols from alum data
d_complier <- left_join(alum_data , open_data)

d_complier[.N]
d_complier[!is.na(status),.N]
print(colnames(d_complier))

#Reading in bounced emails

d_bounce <- fread('data/bounced_emails.csv')
colnames(d_bounce)

head(d_bounce)
setnames(d_bounce, "V2", "Email")

d_bounce[,status := "BOUNCE"]

head(d_bounce)

head(d_complier)

#joining the compliers and the bounced email data
final_data <-left_join(d_complier , d_bounce , by = "Email")

final_data[,.N]

head(final_data)



```

```{r Sanith Check} 


dubios_data_read_bounce <- final_data[(status.y == "BOUNCE") & (status.x == TRUE), ]
#storing
write.csv(dubios_data_read_bounce,"read_bounced.csv",row.names = TRUE)

```


There seems to be something wrng with the `dubious data` as the status can not be `read` and `Bounced` at the same time.
We'll remove these `r dubios_data_read_bounce[,.N]` values from complier count and label them as `Bounced`

```{r updating the status to bounce for these dubious recs }

#https://stackoverflow.com/questions/28650957/update-a-value-in-one-column-based-on-criteria-in-other-columns

class (dubios_data_read_bounce)

colnames(dubios_data_read_bounce)
#we'll update status .x to False

setDT(final_data)[(status.y == "BOUNCE") & (status.x == TRUE), status.x := FALSE]
#verifying the update
final_data[(status.y == "BOUNCE") & (status.x == TRUE), ]

#updating the "Bounce status to TRUE for logical operation"
setDT(final_data)[(status.y == "BOUNCE") , status.y := TRUE]

colnames(final_data)
#updating the names for easier verification
setnames(final_data, "status.y", "Bounce_status")
setnames(final_data, "status.x", "read_status")
colnames(final_data)



```
Based on our final dataset , our next step is to create a final column in our data frame that reflects the status of the emails that are sent.
  - There could be 3 possible status on an email - 'Read, Bounced,Sent'
  - We've already marked as Bounced those emails for which both `Bounced` and `Open` was True
  




```{r}
 final_data<- final_data %>% mutate(final_status = case_when(
    Bounce_status== TRUE ~ "B",
    read_status== TRUE ~ "O"
    
  ))
  

```



```{r}
table(final_data[,final_status])
write.csv(final_data,"final_data_dk.csv",row.names = TRUE)


```

```{r}
table(final_data[, final_status])


```




### Test data
```{r test data}
d1 <- head(d_combined_non_zero)



d_test <- data.table(Email = rep("devesh1412@gmail.com",3),
                     `Merge status` = rep("EMAIL_SENT",3))

d_test2<- data.table(Email = c("devesh1412@gmail.com","dkhandelwal@nvidia.com"),
                     `Merge status` = c(rep("EMAIL_SENT",2)))


d_test3 <- data.table(Email = "dkhandelwal@nvidia.com",`Merge status` = "EMAIL_OPENED")

d_test4 <- data.table(Email = c("dkhandelwal@nvidia.com","dkh@nv.om"),`Merge status` = c("EMAIL_OPENED","EMAIL_BOUNED"))
d_test5 <- data.table(Email = "devesh1412@gmail.com",`Merge status` = "EMAIL_OPENED")
d_test6 <- data.table(Email = "dkhandelwal@nvidia.com",`Merge status` = "EMAIL_CLICKED")
d_test7 <- data.table(Email = "devesh1412@gmail.com",`Merge status` = "EMAIL_CLICKED")
d_test8 <- data.table(Email = "dkhandelwal@yahoo.com",`Merge status` = "EMAIL_SENT")



d_test_new <- rbind(d_test,d_test2,d_test3,d_test4,d_test5,d_test6,d_test7,d_test8)


d_test_new

d_temp <- d_test_new


#d_test_new[`Merge Status`!= "EMAIL_OPENED" ,]

d2 <- setDT(d_temp)[`Merge status`== "EMAIL_OPENED" ,`Emails_status`:= (as.logical(1))]

d2

d2

d2 <- setDT(d_temp)[is.na(`Emails_status`) ,`Emails_status`:= (as.logical(0))]

d2

d2

var_type <- d2[Email == "devesh1412@gmail.com",as.logical(`Merge status`)]


as.logical(d2$`Merge Status`)


class(var_type)

logical_op <- function(col){
    return ((any(col)))
  }



d3 <- d2 %>%  group_by(Email) %>% summarise(status = any(as.logical(`Emails_status`)))


d3

  
  
d3_func_test <- final_mail_status(d_test_new, "EMAIL_OPENED")


d_test_new


d3_func_test



d_test_new

d3_func_test_1 <- final_mail_status(d_test_new, "EMAIL_CLICKED")

d3_func_test_1



d_test_new

test_col <- colnames(d_test_new)

"Merge status" %in%  test_col 



```
